D3064R0: How to Avoid OOTA Without Really Trying

### 2024-03-04 ootavacuous.tex
Updates for the Abstract
The abstract is rather old and in need of updating.  These changes will
put it more in tune with the content of the paper.

Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

\begin{abstract}
	The out-of-thin-air (OOTA) properties of the specification
	of \co{memory_order_relaxed} have resulted in considerable
	consternation over the years.
	Attempts to create memory models that rule out OOTA behaviors
	have either been non-executable, complex, or unloved by C++
	implementers, and this paper summarizes some of the challenges.
	But at the same time, we know of no instances of OOTA behavior
	in real C++ implementations.

	This paper refactors the commonly used condition for classifying
	OOTA behavior, which permits bringing to bear constraints from
	long-standing laws of physics and from hardware systems.
	In this initial work, we focus on shared-memory programs running
	on C++ implementations based on compilers and computing hardware
	(including CPUs and GPGPUs).
	We present informal arguments that these constraints and
	limitations prevent OOTA cycles containing only volatile atomics
	from occurring in undefined-behavior-free C++ programs running on
	correctly constructed C++ compiler-based implementations that are
	restricted to single-threaded analysis in our current universe.
	We extend this work to nonvolatile atomics by defining
	quasi-volatile behavior that we conjecture that usable real-world
	C++ implementations must adhere to.
\end{abstract}

>>>>

\begin{abstract}
	The out-of-thin-air (OOTA) properties
	of \co{memory_order_relaxed} in the C++ memory model
	have caused considerable consternation over the years.
	Attempts to create memory models that rule out OOTA behaviors
	have been either non-executable, complex, or unloved by C++
	implementers
	(and this paper describes some of the challenges of coming to
	grips with the key notion of semantic dependency).
	But at the same time, we know of no instances of OOTA behavior
	in real C++ implementations.

	We focus on shared-memory programs and
	C++ implementations based on traditional compilers and
	computing hardware, including CPUs and GPGPUs.
	This context permits us to consider the detailed relations
	between source code and machine code required by the restricted
	nature of volatile atomic accesses.
	The OOTA problem is much more tractable at the hardware level
	than at the source level, thanks to existing formal hardware
	models, and
	we show that the constraints in these models prevent OOTA cycles
	involving only volatile atomics from occurring in
	undefined-behavior-free C++ programs (under these implementations).
	We also extend this work to nonvolatile atomics by defining
	``quasi-volatile'' behavior that we expect C++ implementations
	will adhere to if they perform single-thread analysis.
\end{abstract}
